### 2020-04-25 generated/paper_EN.Rmd
Add papers

```{r similar-samples-2, echo=FALSE}
# Tabela - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).

@@

```{r join-90-91-100, echo=F, eval=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

@@

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
# Table - First 6 samples after reorganization/EN>
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}
# Total samples/EN>
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables
To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
```{r new-features-2, echo=F}
# Adding New features/EN>
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```
```{r new-features-table, echo=F}
# Table - New features/EN>
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

@@

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables/EN>
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables/EN>
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

@@

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age/EN>
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
                "14", "28", "56", "100",
                "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

@@

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age/EN>
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
mix_dat$day <- factor(mix_dat$day)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features/EN>
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

@@

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
# Figure - Variables distribution/EN>

>>>>

```{r similar-samples-2, echo=FALSE}
# Tabela - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.

@@

```{r join-90-91-100, echo=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

@@

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
# Table - First 6 samples after reorganization
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}
# Total samples
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables
To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
```{r new-features-2, echo=F}
# Adding New features
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```
```{r new-features-table, echo=F}
# Table - New features
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

@@

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

@@

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
                "14", "28", "56", "100",
                "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
# Figure - Correlation over time
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

@@

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
mix_dat$day <- factor(mix_dat$day, levels=lvls)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

@@

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
# Figure - Variables distribution
